---
title: "Victimology_factsheet"
author: "Sofia F. M. Schmidt-Kallesøe"
date: "6/2/2019"
output: html_document
---

```{r}

setwd("~/Documents/Amsterdam 19/Victimology")

library(haven)
library(dplyr)
library(data.table)
library(foreign)
library(ggplot2)
library(pastecs)

#JW_data <- read_sav("Just World  data.sav")
#View(JW_data)

#Creating .csv file 
#loc1 = read.spss("~/Documents/Amsterdam 19/Victimology/Just World  data.sav", to.data.frame=TRUE)
#write.csv(loc1, file="~/Documents/Amsterdam 19/Victimology/loc1.csv")

loc = read.csv("~/Documents/Amsterdam 19/Victimology/loc1.csv")

```


Subsetting locus of control questions, victimisation experiences, and demographics 
```{r}

loc = select(loc, 2, 23:63)

```

Renaming variables
```{r}

colnames(loc) = c("ID", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y", "z", "aa", "bb", "cc", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "Gender", "Age", "Income", "Education", "Urban_or_rural")

```

```{r}

#Creating data frame with victimisation questions and demographics to merge with the locus of control score points data frame 
demographics = subset(loc, select = c(ID, V3, V4, V5, V6, V7, V8, V9, Gender, Age, Income, Education, Urban_or_rural))

```


Locus of control score 
```{r}

locscoredf <- loc %>% 
  select(b,c,d,e,f,g,i,j,k,l,m,o,p,q,r,t,u,v,w,y,z,bb,cc) %>%  #select relevant columns
  mutate(b = if_else(b == 1, true = 1, false = 0),
         c = if_else(c == 2, true = 1, false = 0),
         d = if_else(d == 2, true = 1, false = 0),
         e = if_else(e == 2, true = 1, false = 0),
         f = if_else(f == 1, true = 1, false = 0),
         g = if_else(g == 1, true = 1, false = 0),
         i = if_else(i == 1, true = 1, false = 0),
         j = if_else(j == 2, true = 1, false = 0),
         k = if_else(k == 2, true = 1, false = 0),
         l = if_else(l == 2, true = 1, false = 0),
         m = if_else(m == 2, true = 1, false = 0),
         o = if_else(o == 2, true = 1, false = 0),
         p = if_else(p == 1, true = 1, false = 0),
         q = if_else(q == 1, true = 1, false = 0),
         r = if_else(r == 1, true = 1, false = 0),
         t = if_else(t == 1, true = 1, false = 0),
         u = if_else(u == 1, true = 1, false = 0),
         v = if_else(v == 2, true = 1, false = 0),
         w = if_else(w == 1, true = 1, false = 0),
         y = if_else(y == 1, true = 1, false = 0),
         z = if_else(z == 2, true = 1, false = 0),
         bb = if_else(bb == 2, true = 1, false = 0),
         cc = if_else(cc == 1, true = 1, false = 0)
         ) %>% 
  mutate(points = rowSums(.)) %>% #the "." if a refference to the dataframe which is being passed on by the pipe " %>% "
  cbind(select(loc, ID)) #binds the passed df with the id columns from the testdf

#Fixing the order of the columns
locscoredf <- locscoredf %>% select(ID, points, everything())

#Bind demographics+victimisation questions df and locscoredf together
locdf = merge(locscoredf, demographics, by = c("ID"))


```

Changing all answers that are not no or yes to NA (to conduct t-tests, which requires a factor with only two levels)
```{r}

locdf[locdf=="Do not want to tell"] <- NA
locdf[locdf=="Do not know"] <- NA

```


Creating a column for the added points for each participant
```{r}

#dataframe = within(locdf, {
#    loc = ifelse(points < "12"| points > "12", 1, 2)
# })

locdf$loc = if_else(locdf$points < 13, true = 1, false = 2)

```

Checking for normality before conducting tests 
```{r}

ggplot(locdf, aes(points)) + 
  geom_histogram(aes(y=..density..), colour = "darkblue", fill = "lightblue", binwidth = 3) + 
  stat_function(fun = dnorm, args = list(mean(locdf$points), sd(locdf$points)), size = 1) + 
  labs(x = "Points", y = "Density") + 
  ggtitle("Points") +
  theme_minimal()

qplot(sample = locdf$points)

round(stat.desc(locdf$points, basic = FALSE, norm = TRUE),4)

```


V3 variable - Have you ever been a victim of an offence?
```{r}

#Boxplot of the two groups (excluding NAs)
ggplot(subset(locdf, !is.na(V3)), aes(x=V3, y=points)) + 
  geom_boxplot(width = 0.3, fill = "cornflowerblue") +
  labs(y = "Locus of Control score", x = "Have you ever been victimised?") +
  theme_minimal()

#Descriptive statistics
by(locdf$points, locdf$V3, stat.desc, basic = FALSE, norm = TRUE)


#Independent t-test 
V3.ttest = t.test(points ~ V3, data = locdf)
V3.ttest


#Calculating effect size 
t = V3.ttest$statistic[[1]] #Creating a variable t that contains the value of t 

df = V3.ttest$parameter[[1]] #Creating a variable df containing the degrees of freedom

r = sqrt(t^2/(t^2+df)) #Calculating r 
round(r, 3)

```

On average, participants had a higher LOC score when having been victimised (M = 16.3, SE = .42), than when not having been victimised (M = 12.96, SE = 0.39). This difference was significant t(119.34) = 5.86, p < .01 and had an effect of r = .47. 



V4 variable - Have you been a victim of a violent offence in the past year?
```{r}

#Boxplot of the two groups (excluding NAs)
ggplot(subset(locdf, !is.na(V4)), aes(x=V4, y=points)) + 
  geom_boxplot(width = 0.1) +
  labs(x = "Locus of Control score", y = "V4. Have you been a victim of a violent offence in the past year?") +
  theme_minimal()


#Descriptive statistics
by(locdf$points, locdf$V4, stat.desc, basic = FALSE, norm = TRUE)


#Independent t-test 
V4.ttest = t.test(points ~ V4, data = locdf)
V4.ttest


#Calculating effect size 
t = V4.ttest$statistic[[1]] #Creating a variable t that contains the value of t 

df = V4.ttest$parameter[[1]] #Creating a variable df containing the degrees of freedom

r = sqrt(t^2/(t^2+df)) #Calculating r 
round(r, 3) 

```

On average, participants had a higher LOC score when having been victimised within the last year (M = 16.06, SE = .61), than when not having been victimised (M = 13.59, SE = 0.34). This difference was significant t(27.2) = 3.54, p < .01 and had an effect of r = .56. 



V5 variable - Have you been a victim of a sexual offence in the past year?
```{r}

#Boxplot of the two groups (excluding NAs)
ggplot(subset(locdf, !is.na(V5)), aes(x=V5, y=points)) + 
  geom_boxplot(width = 0.3, fill = "cornflowerblue") +
  labs(y = "Locus of Control score", x = "Sexual offense - Past year") +
  theme_minimal()

#Descriptive statistics
by(locdf$points, locdf$V5, stat.desc, basic = FALSE, norm = TRUE)


#Independent t-test 
V5.ttest = t.test(points ~ V5, data = locdf)
V5.ttest


#Calculating effect size 
t = V5.ttest$statistic[[1]] #Creating a variable t that contains the value of t 

df = V5.ttest$parameter[[1]] #Creating a variable df containing the degrees of freedom

r = sqrt(t^2/(t^2+df)) #Calculating r 
round(r, 3) 

```

On average, participants had a higher LOC score when having been sexually assaulted within the last year (M = 19.71, SE = .42), than when they hadn't (M = 14, SE = 0.31). This difference was significant t(13.84) = 10.99, p < .01 and had an effect of r = .95. 




V6 variable - Have you been a victim of a theft in the past year?
```{r}

#Boxplot of the two groups (excluding NAs)
ggplot(subset(locdf, !is.na(V6)), aes(x=V6, y=points)) + 
  geom_boxplot() +
  labs(x = "Locus of Control score", y = "V6. Have you been a victim of a theft in the past year?") +
  theme_minimal()

#Descriptive statistics
by(locdf$points, locdf$V6, stat.desc, basic = FALSE, norm = TRUE)


#Independent t-test 
V6.ttest = t.test(points ~ V6, data = locdf)
V6.ttest


#Calculating effect size 
t = V6.ttest$statistic[[1]] #Creating a variable t that contains the value of t 

df = V6.ttest$parameter[[1]] #Creating a variable df containing the degrees of freedom

r = sqrt(t^2/(t^2+df)) #Calculating r 
round(r, 3) 

```

On average, participants had a higher LOC score when having been a victim of theft within the last year (M = 16.5, SE = .48), than when they hadn't (M = 13.26, SE = 0.35). This difference was significant t(58.94) = 5.43, p < .01 and had an effect of r = .58. 




V7 variable - Have you been a victim of a robbery in the past year?
```{r}

#Boxplot of the two groups (excluding NAs)
ggplot(subset(locdf, !is.na(V7)), aes(x=V7, y=points)) + 
  geom_boxplot() +
  labs(x = "Locus of Control score", y = "V7. Have you been a victim of a robbery in the past year?") +
  theme_minimal()

#Descriptive statistics
by(locdf$points, locdf$V7, stat.desc, basic = FALSE, norm = TRUE)


#Independent t-test 
V7.ttest = t.test(points ~ V7, data = locdf)
V7.ttest


#Calculating effect size 
t = V7.ttest$statistic[[1]] #Creating a variable t that contains the value of t 

df = V7.ttest$parameter[[1]] #Creating a variable df containing the degrees of freedom

r = sqrt(t^2/(t^2+df)) #Calculating r 
round(r, 3) 

```

On average, participants had a higher LOC score when having been a victim of robbery within the last year (M = 16.33, SE = .80), than when they hadn't (M = 13.53, SE = 0.33). This difference was significant t(23.17) = 3.22, p < .01 and had an effect of r = .56. 


V8 variable - Have you been a victim of vandalism in  the past year?
```{r}

#Boxplot of the two groups (excluding NAs)
ggplot(subset(locdf, !is.na(V8)), aes(x=V8, y=points)) + 
  geom_boxplot(width = 0.3, fill = "cornflowerblue") +
  labs(y = "Locus of Control score", x = "Vandalism - past year") +
  theme_minimal()

#Descriptive statistics
by(locdf$points, locdf$V8, stat.desc, basic = FALSE, norm = TRUE)


#Independent t-test 
V8.ttest = t.test(points ~ V8, data = locdf)
V8.ttest


#Calculating effect size 
t = V8.ttest$statistic[[1]] #Creating a variable t that contains the value of t 

df = V8.ttest$parameter[[1]] #Creating a variable df containing the degrees of freedom

r = sqrt(t^2/(t^2+df)) #Calculating r 
round(r, 3) 

```

On average, participants had a higher LOC score when having been a victim of vandalism within the last year (M = 18, SE = 1.56), than when they hadn't (M = 13.49, SE = 0.33). This difference was significant t(7.62) = 2.83, p < .05 and had an effect of r = .72. 



V9 variable - To your knowledge has anyone of your close friends and/or family been a victim of an offence in the past year? 
```{r}

#Boxplot of the two groups (excluding NAs)
ggplot(subset(locdf, !is.na(V9)), aes(x=V9, y=points)) + 
  geom_boxplot() +
  labs(x = "Locus of Control score", y = "V9. To your knowledge has anyone of your close friends 
       and/or family been a victim of an offence in the past year?") +
  theme_minimal()

#Descriptive statistics
by(locdf$points, locdf$V9, stat.desc, basic = FALSE, norm = TRUE)


#Independent t-test 
V9.ttest = t.test(points ~ V9, data = locdf)
V9.ttest


#Calculating effect size 
t = V9.ttest$statistic[[1]] #Creating a variable t that contains the value of t 

df = V9.ttest$parameter[[1]] #Creating a variable df containing the degrees of freedom

r = sqrt(t^2/(t^2+df)) #Calculating r 
round(r, 3) 

```

On average, participants had a higher LOC score when having been a victim of theft within the last year (M = 14.24, SE = .41), than when they hadn't (M = 13.18, SE = 0.52). This difference was not significant t(111.11) = 1.6, p = 0.11 and only had a small effect of r = .15. 



More descriptive statistics 
```{r}

#Percentage of male victims 
mv = subset(locdf, Gender == "male")
summary(mv)
#Yes=21, no=20, NA=9

21/41*100 
#51.22% of male participants have been victimised


#Percentage of female victims 
fv = subset(locdf, Gender == "female")
summary(fv)
#Yes=29, no=65, NA=6

29/94*100
#30.85% of female participants have been victimised 


#Percentage of victims 
summary(locdf)
50/135*100
#37.04% of participants have been victimised 


#Percentage of victims with an internal LoC
iv = subset(locdf, loc == "1")
summary(iv)
#Yes=5, no=37, NA=8

5/42*100
#11.9% of participants with an internal locus of control have been victimised


#Percentage of victims with an external LoC 
ev = subset(locdf, loc == "2")
summary(ev)
#Yes=45, no=48, NA=7

45/93*100
#48.39% of participants with an external locus of control have been victimised

```


Gender differences in victimisation
```{r}

#Seeing whether gender predicts victimisation
locdf$Gender<-relevel(locdf$Gender, "female")
locdf$V3<-factor(locdf$V3, levels = c("no", "Yes"))
as.numeric(locdf$Gender)

gv_model = glm(V3 ~ Gender, data = locdf, family = binomial())
summary(gv_model)

#Gender was a significant predictor of being victimised, b = .86, z = 2.23, p < .05. Males are more likely to victimised than females. 


#Testing model significance
modelChi <- gv_model$null.deviance - gv_model$deviance
modelChi

chidf <- gv_model$df.null - gv_model$df.residual

chisq.prob <- 1 - pchisq(modelChi, chidf)
chisq.prob

#The p-value is 0.03, and is thus significant meaning we can reject the null hypothesis that the model is not better at predicting than chance. 

#Finding odds ratio and CI
exp(gv_model$coefficients)

exp(confint(gv_model))

```

